Implicit yaddayadda

### 2011-09-20 papers/PLDI2012/implicitization.tex
Explication de la technique combinée en cours...

\usepackage{amsmath,amsfonts}
\usepackage{dmnatbib}
\usepackage[mathscr]{euscript}

@@

\section{Guided static analysis}
\label{sec:guided_multigraph}
\emph{Guided static analysis} was proposed by \citet{DBLP:conf/sas/GopanR07} as an improved over classical upward Kleene iterations with widening.
Consider the following program:
\lstinputlisting{gopan_reps.c}

@@

>>>>

\usepackage{amsmath,amsfonts}
\usepackage{dmnatbib}
\usepackage[mathscr]{euscript}
\usepackage{algorithmic,algorithm}

@@

\subsection{Guided static analysis}
\label{sec:guided_multigraph}
\emph{Guided static analysis} was proposed by \citet{DBLP:conf/sas/GopanR07} as an improved over classical upward Kleene iterations with widening.
Consider the following program:
\lstinputlisting{gopan_reps.c}

@@

\section{Guided analysis over the paths}
\label{sec:combined}

Guided static analysis \cite{DBLP:conf/sas/GopanR07} applies to the transition
graph of the program. One could instead apply the same technique to the implicit
multigraph from \cite{Monniaux_Gonnord_SAS11}, thus avoiding control flow merges with
unfeasible paths.

Let $P$ be the set of control points in the transition graph, 
$P_W \subseteq P$ the set of widening points
such that removing the points in $P_W$ gives an acyclic graph.
One can choose a set $P_R$ such that $P_W \subseteq P_R \subseteq P$.
The combination of these two techniques considers an ascending sequence of
subsets of the paths between two points in $P_R$.

The set of paths is kept implicit by an SMT formula $\rho$ expressing
the semantics of the program, assuming that the transition semantics can be
expressed within a decidable theory.

This formula contains Boolean \emph{reachability predicates} $b_i$ for each
control points $p_i \notin P_R$, $b_i^s$ and $b_i^d$ for each $p_i \in P_R$, so
that a path 
$p_{i_1} \rightarrow p_{i_2} \rightarrow \dots \rightarrow p_{i_n}$ 
between two points $p_{i_1}, p_{i_n} \in P_R$ 
can easily be expressed as the
cunjunction $b_{i_1}^s \wedge \bigwedge_{2 \leq k < n} b_{i_k} \wedge b_{i_n}^d$.

The current working subset of paths is stored using a compact representation,
such as Binary Decision Diagrams.

\subsection{Ascending iterations}
\label{subsec:ascending}

The algorithm for computing an inductive invariant over a subgraph is the Path
Focused algorithm from \citet{Monniaux_Gonnord_SAS11}.

In order to find which path to focus on, we construct an SMT formula $f(p_i)$, whose
model when satisfiable is a path that starts in $p_i$, goes to a successor $p_j
\in P_R$ of $p_i$, such that the image of $X_{p_i}^s$ by the path transformation
is not included in the actual $X_{p_j}^s$.
Intuitively, such a path makes the abstract value $X_{p_j}^s$ grow, and thus is
an interesting path to focus on. When the formula becomes unsatisfiable, it
means the analysis of $p_i$ is finished.

If we note $Succ(i)$ the set of indices $j$ such that $p_j \in P_R$ is a
successor of $p_i$ in the expanded multigraph, and $X_i^s$ the abstract value
associated to $p_i$ :
$$f(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge x \in X_i^s \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
x \notin X_j^s)$$

Since we do not work on the entire transition graph but on a subset of it, we
have to cunjunct the formula $f(p_i)$ with the actual set of working paths,
noted $P$, expressed as a boolean formula, where the boolean variables are the
\emph{reachability predicates} of the control points. We can easily construct
this formula from the binary decision diagram using dynamic programming, and
avoiding an exponentially sized formula. In other words, we force the SMT solver
to give us a path which is in $P$.

\subsection{Adding new paths}
\label{subsec:addingpaths}

Our technique computes the fixpoint iterations on a ascending sequence of
subgraphs, until the complete graph is reached.

When the analysis of a subgraph is finished, meaning that the abstract values
for each control points has converged to an inductive invariant for this subgraph,
the next subgraph to work on has to be computed.

This new subgraph is the union of the paths of the previous one with a set
$P'$ of new paths that become feasible regarding the current abstract values.
$P'$ is progressively constructed during the analysis of the subgraph.

We distinguish two phases in the analysis of a point $p_i \in P_R$:
\begin{enumerate}
\item We search for paths starting at point $p_i$, such that these paths are in
$P$, i.e are included in the working subgraph. Each time we find a path, we
update the abstract value of the destination point of the path. This is the
phase explained in \ref{subsec:ascending}.
\item  When the precedent phase terminates, there is no more paths in
$P$, starting at $p_i$, 
that make the abstract values of the successors grow (otherwise, the SMT
solver would not have answered ``\emph{unsat}''). However, there may
exist such paths in the entire multigraph, that are not in $P$. The second phase
computes these paths and adds them to $P'$.
\end{enumerate}

It is not necessary to add into $P'$ all the new feasible paths. We just have
to add a new path into $P'$ until no more path can make the invariant grow,
as proposed by Algorithm \ref{algo:computepaths}. For that, we use another
abstract value associated to a control point $p_j$, noted $X_j^d$, which is
different from $X_j^s$.

\begin{algorithm}
	\caption{ComputeNewPaths}
	\label{algo:computepaths}
	\begin{algorithmic}[1] 
	\input{algo/compute-paths}
	\end{algorithmic}
\end{algorithm}

\subsection{Algorithm}

The algorithm for Guided static analysis on the implicit multigraph is described
in Algorithm \ref{algo:combined}.

We maintain a set of active control points $A$, containing the points in $P_R$
that still need to be processed. We apply the two previously described phases to
the elements of $A$ until $A$ is empty. When $A$ is empty, it means the
invariant computation of the subgraph is finished. As proposed by
\citet{DBLP:conf/sas/GopanR07}, we can do some narrowing iterations, and we
restart the analysis with a new subgraph if needed.
\begin{algorithm}
	\caption{Guided static analysis on implicit multigraph}
	\label{algo:combined}
	\begin{algorithmic}[1] 
	\input{algo/algo-combined}
	\end{algorithmic}
\end{algorithm}
